---
title: "Logistic Regression"
author: "Barboza-Salerno"
---

# Problem Set 5

Send me the markdown file by 3/1/2024 \@ 11:59. The subject line should say "Problem Set 5" or it will receive no credit!

In this problem set you are going to examine the 2021 [National Survey of Child Health (NSCH)](https://www.census.gov/programs-surveys/nsch/data/datasets.2021.html#list-tab-491554181). First, navigate to the website and download the file called `Revised SAS data file`. The file is a zip file so you must unzip it before attempting to use it.

## PhD students use R

### Open the file in R using the sas7bdat package.

::: callout-note
The foreign package is typically used to open files with extensions associated with STATA, SPSS, and SAS, however, this one works fine with different versions and so its easier.
:::

```{r, echo=T, message=FALSE, results="hide"}
library(sas7bdat)
nchs <- read.sas7bdat("C:/Users/barboza-salerno.1/Downloads/nsch_2021_topical_SAS/nsch_2021e_topical.sas7bdat")
```

```{r}
library(labelled)
var_label(nchs$BULLIED_R) <- "Bullied in last 12 months"
```

```{r}
table(nchs$BULLIED_R)
get_variable_labels(nchs$BULLIED_R)
```
```{r}
nchs <- nchs %>% dplyr::rename("Bullied" = "BULLIED_R")
```

```{r}
table(nchs$Bullied)
```

You will definitely want to use the [codebook](https://www.census.gov/data-tools/demo/uccb/nschdict) to get a sense of the variable definitions.

### Research Question

What individual and contextual factors are associated with brain injury among young children?

### Dependent variable

You are to use the dependent variable `CONFIRMINJURY.` Recode the variable into a binary variable 0/1.

::: callout-note
The dependent variable is the variable associated with a confirmed injury (i.e., from a medical provider). However, only persons who reported an injury are asked this question. So, you need to recode accordingly.
:::

### Independent variable

Select two individual characteristics (e.g., age of child) and two 'contextual' characteristics (e.g., home enviornment).

### Perform an appropriate multivariable analysis to adress the RQ

### Presentation

-   Create a table that presents your results similar to the one I use in [my example](https://phhbhp7534.netlify.app/example/logistic). Make sure you have the odds ratios in the table.

-   Select an interesting result and create a graphical presentation to illustrate the result

### Interpret the results

Interpret the odds ratios for each variable and whether they are statistically significant.

## Master's students use R, JASP or JAMOVI, your choice
