---
title: "Moderation Analysis"
author: "Barboza-Salerno"
---

```{r}
library(foreign) # this package allows you to open SAS, SPSS and STATA files
library(dplyr)
library(DiagrammeR)
```
# Moderation in R

The dataset used in this example is located [here](https://github.com/bigdataforsocialjustice/PHHBHP7534/blob/main/moderation.sav)

## Introduction
Moderation analysis examines whether the relationship between two variables, $X$ and $Y$ depends on the effect of another variable, denoted $W$. We refer to the variable $W$ as the moderator. Moderators can be categorical or continuous. Gender is a typical categorical variable that is often used as a moderator. Moderators can also be continuous, such as the effect of post-traumatic symptoms on delinquent behavior depends on depressive symptoms. To illustrate moderation I will focus on whether depressive symptoms moderate the impact of post-traumatic stress on delinquent behaviors among youth who have had contact with the child welfare system.

:::{.callout-note}
This is a great analysis to perform when you are interested in examining whether an intervention has a different effect for different groups of people: does the intervention work similarly for different groups of people. Moderation can also be used to examine whether a variable is protective in reducing the effect of one variable on another
:::

It is very useful to visualize the relationships between the variables, which is something that **no other software allows you to do besides R, at least do it easily and beautifully.** For more information about the visualizations below see this [website](https://rich-iannone.github.io/DiagrammeR/articles/graphviz-mermaid.html). You should be able to copy and paste the code to create your own visualizations.

```{r}
# global setting
grViz("digraph flowchart {
graph [layout = neato, overlap = TRUE, rankdir = TB, label = 'Figure 1. Moderation (Hayes Model 1 )']
# node definitions with label text and position
node [fontname = 'Times New Roman', shape = rectangle]        
tab1 [label = 'PTSS', pos = '1,1!']
tab2 [label = 'DELINQ', pos = '5,1!']
tab3 [label = 'DEPRESS', pos = '3,2!']
tab4 [shape = point, pos = '3,0.97!', color = white]
# edge definitions with the node IDs
tab1 -> tab2;
tab3 -> tab4  }")
```

Figure 1 is the theoretical model of moderation. The model describes how the effect of a focal predictor, $X$ on the outcome, or dependent variable, $Y$, depends on levels of the moderator, $W$.

Moderation is simply an interaction between the focal predictor $X$ and the moderator, $W$. If, after you include the interaction term in the model, which is simply done by multiplying $X$ by $W$, the coefficient on the interaction term is significant, then we say that the effect of $X$ on $Y$ is moderated by $W$.

## Moderation Example
The data are from the National Survey of Child and Adolescent Well-Being and can be accessed from the [National Data Archive on Child Abuse and Neglect](https://www.ndacan.acf.hhs.gov/datasets/datasets-list.cfm). I am using only variables from Wave I for this illustration. The variables I am using are

|Variable | Measure |
|---------|:-----|------:|:------:|
|CIDI |  Depressive symptoms |
|TRA | PTS symptoms |
|BCDEL | Delinquent behavior |
|SEXAB | Experienced sexual abuse|

```{r}
setwd("C:/Users/barboza-salerno.1/Downloads/")
moderation_example = tibble::as_tibble(read.spss("moderation.sav", use.value.labels = F, to.data.frame = T))
moderation_example
```
To facilitate interpretation we mean center the variables first before including them in the analysis. The easiest way to do that is shown below.

```{r}
moderation_example$tra1_mean <- moderation_example$tra1 - mean(moderation_example$tra1, na.rm = T)

moderation_example$cidi1_mean <- moderation_example$cidi1 - mean(moderation_example$cidi1, na.rm = T)
```


```{r}
moderation.lm = lm(bcdel1 ~ tra1_mean * cidi1_mean, moderation_example)
summary(moderation.lm)
```

The significant interaction tra1:cidi1 shows that the effect of post-traumatic stress symptoms on engagement in delinquent behavior is moderated by depressive symptoms.

#### PROCESS
The process macro is available in SPSS, SAS and R. Since we are using R, we need to install the appropriate library. To make matters very confusing, there is a package called `processr` and one called `processR`

Neither package is available on the CRAN network. So, we need to install it from the creators Github website. To install type `if(!require(devtools)) install.packages("devtools")`
`devtools::install_github("cardiomoon/processR")`

Note you must install the package `devtools` before running this command either from the file menu or by typing `install.packages("devtools")`

```{r}
library(processR)
```

```{r}
fit <- lm(bcdel1 ~ tra1_mean * cidi1_mean, moderation_example)
```

### 
```{r}
labels=list(X="PTSS",W="DEP",Y="DEL")
pmacroModel(1,labels=labels)
```

```{r}
statisticalDiagram(1,labels=labels)
```

```{r}
modelsSummaryTable(list(fit),labels=labels)
```

```{r}
condPlot(fit,mode=3,xpos=0.5)
```
The plot shows that the effect of PTSS (*X*) on Delinquent Behavior (*Y*) is (significantly) positive for people with lower levels of depressive symptoms and (significantly) negative for people high in depressive symptoms. 


```{r}
condPlot(fit,mode=2,xpos=0.5)
```


```{r}
jnPlot(fit,plot=FALSE)
```
The Johnson-Neyman technique enables us to identify regions of significance. This is where the conditional effect of depressive symptoms (cidi1) alters the relationship between post-traumatic stress symptoms and delinquent behavior.

The boundaries of the Johnson-Neyman regions of significance are defined by the values -9.642 and 15.311. This is where the conditional effect of depressive symptoms (cidi1) *significantly* and negatively alters the relationship between post-traumatic stress symptoms and delinquent behavior.
