---
title: "Key Steps in the research process"
date: "2024-01-09"
editor: 
  markdown: 
    wrap: 72
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sHcnk9KcoQVzDywxWRzVWHC"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "S4OmWxc3h3Y", "Lecture Slides", "1", TRUE
 # "DGbLEgL4g6Y", "Drawing lines", "drawing-lines", TRUE,
#  "n-rV1TIoIgw", "Lines, Greek, and regression", "lines-greek-regression", FALSE,
 # "RYKsrTKWCR4", "Null worlds and statistical significance", "significance", FALSE
)
```

::: callout-note
Course objective(s) for this module: *Describe and execute the key steps
in designing, conducting, and evaluating research in health behavior and
health promotion.*
:::

## Readings

### Today's Assignments

-   The [syllabus](/syllabus.qmd), [content](/content/index.qmd),
    [examples](/example/index.qmd), and
    [assignments](/assignment/index.qmd) pages for this class
-   {{< fa book >}} [Chapter 1](/files/rmhp/ch1.pdf), p. 3-22 in
    *Research Methods in Health Promotion* [@crosby_research_2006]

### Your tasks

-   [ ] Download and install [R/Rstudio](/resource/install.qmd)
-   [ ] Download and install [Jamovi Desktop](https://www.jamovi.org)
-   [ ] Download and install [JASP](https://jasp-stats.org/download/)
-   [ ] Download and install [Zotero](/resource/zotero.qmd)
-   [ ] Connect to the course
    [Slack](https://app.slack.com/client/T06BQF8BPP0/C06BZJ1HHS8)
-   [ ] [Optional] Download the [BibTex file](/resource/citations.qmd) of the course readings
-   [ ] (optional) Download and install
    [Obsidian](https://obsidian.md/download)

## Lecture Notes

:::{.callout-note}
The slides will only be made available immediately after class. Only then will the 'Page Not Found' error message go away.
:::

The slides for today's lesson are available online as an HTML file. Use
the buttons below to open the slides either as an interactive website or
as a static PDF (for printing or storing for later). You can also click
in the slides below and navigate through them with your left and right
arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/day1")
slide_tabs(slide_details, "/slides/day1.html")
```

::: callout-tip
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> +
<kbd>/</kbd>) while going through the slides, you can see a list of
special slide-specific commands. Use right arrow <kbd>-\></kbd> to
scroll through slides
:::

## In-class Materials

Here are all the materials we'll use in class:

-   **Intro to R** [webpage](/slides/intro-to-r.html)
-   **Intro to R** [data and markdown](/slides/intro to r.zip)
-   **Intro to Plotting with R** [webpage](/slides/plots.html)
-   **Intro to Plotting with R** [data and markdown](/slides/plots.Rmd)

### Software Packages

-   [Add-on Packages &
    Modules](https://bigdataforsocialjustice.github.io/databootcamp/add-on-packages-&-modules.html)
    offer additional functionality and are available for all software
    programs listed above

### 📦 Additional Resources

-   [user guide - jamovi](https://www.jamovi.org/user-manual.html)
-   [jamovi library](https://www.jamovi.org/library.html)
-   [jmv r package - jamovi](https://www.jamovi.org/jmv/)
-   [community resources -
    jamovi](https://www.jamovi.org/community.html)
-   In addition, once jamovi is installed, there is a 'data library'
    available to you. This data library has worked examples for all
    modules that you install.

### Competencies and Knowledge

-   Foundational knowledge: 3, 6
-   Competencies: 9
-   MPH Specializations: 4
-   PhD Specializations: 4

### Student Evaluation Activity for Assessment

-   Exam 1
-   Proposal 1
