---
title: "Variables & Units of Analysis"
date: "2024-01-11"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sERjvBbJGr9mpqvtiLVPzd4"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "S4OmWxc3h3Y", "Lecture Slides", "1", TRUE
 # "DGbLEgL4g6Y", "Drawing lines", "drawing-lines", TRUE,
#  "n-rV1TIoIgw", "Lines, Greek, and regression", "lines-greek-regression", FALSE,
 # "RYKsrTKWCR4", "Null worlds and statistical significance", "significance", FALSE
)
```

::: callout-note
Course objective(s) for this module: *Describe and execute the key steps
in designing, conducting, and evaluating research in health behavior and
health promotion.*
:::

## Readings

- {{< fa book >}} *Chapter 4* Babbie, E. R. (2016). The practice of social research (Fourteenth edition). Cengage Learning.[@babbie_practice_2016]
- {{< fa book >}} Barboza-Salerno, G. E. (2020). [Variability and stability in child maltreatment risk across time and space and its association with neighborhood social & housing vulnerability in New Mexico: A bayesian space-time model.](https://www.sciencedirect.com/science/article/pii/S0145213420301277?casa_token=ooa5R4MswOUAAAAA:HdE66nKYJ0wFfW5caFu5X4mY7EufVEjuw2bVYTkgu8WDRdQB4Qk71F6etvkTkRsbJtPs_3Z7d3Sp) Child Abuse & Neglect, 104, 104472. https://doi.org/10.1016/j.chiabu.2020.104472 [@barboza-salerno_variability_2020]
- {{< fa book >}} Yang, T.-C., Kim, S., Zhao, Y., & Choi, S. E. (2021). [Examining spatial inequality in COVID-19 positivity rates across New York City ZIP codes.](https://www.sciencedirect.com/science/article/pii/S1353829221000708?casa_token=sC5jpoput9sAAAAA:LIrijgDzfcje9BI60uBZOKoKyMrgs6ShPkDWE11M7iNm7vinDkq_GlTzPGZFCDfEZCQtFfPxqoLJ) Health & Place, 69, 102574. https://doi.org/10.1016/j.healthplace.2021.102574 [@yang_examining_2021]
- {{< fa book >}} Moss, J. L., Reiter, P. L., & Brewer, N. T. (2015). [Correlates of human papillomavirus (HPV) vaccine coverage: A state-level analysis.](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4295643/) Sexually Transmitted Diseases, 42(2), 71–75. https://doi.org/10.1097/OLQ.0000000000000225 [@moss_correlates_2015]

## Recommended readings

The following reading is a great resources, and I will incorporate some of the material in the lecture.

- {{< fa book >}} Chapter 4. Singleton Jr, R. (2017). [Approaches to social research.](https://global.oup.com/ushe/product/approaches-to-social-research-9780190614249?cc=us&lang=en&) Oxford University Press. 

### Make sure you have completed these tasks by the end of the week

-   [ ] Download and install [R/Rstudio](/resource/install.qmd)
-   [ ] Download and install [Jamovi Desktop](https://www.jamovi.org)
-   [ ] Download and install [JASP](https://jasp-stats.org/download/)
-   [ ] Download and install [Zotero](/resource/zotero.qmd)
-   [ ] Connect to the course [Slack](https://app.slack.com/client/T06BQF8BPP0/C06BZJ1HHS8)
-   [ ] (optional) Download and install [Obsidian](https://obsidian.md/download)

### Your tasks for today

-   [ ] Think of a research question you want to pursue this semester.
-   [ ] Complete [Problem Set 1](/assignment/01-problem-set.qmd). The [instructions for submission](/assignment/01-problem-set.html#task-3-work-with-r) are located in the assignment 

### Notes

- If you are an MPH student, my suggestion is that you select a research topic that is aligned with a final assessment you need to graduate (e.g., APE)
- If you are a PhD student, make sure you select a question that can be answered within the semester. This means
  - Identifying an already existing dataset. 
  - To get you started, look at the data I compiled last summer for the [data bootcamp](https://bigdataforsocialjustice.github.io/databootcamp/) in Social Work by [clicking here](https://bigdataforsocialjustice.github.io/databootcamp/datasets.html); and
  - Collaborating with myself, your advisor, and/or a colleague; and
  - The question must be approved by me; and
  - The topic is due by next class. If you have no topic by then, I will assign a topic to you.

### In-class activities
- Literature search
- Introduction to [Zotero](/resource/zotero.qmd)
  - Google scholar/Zotero integration
  - Word/Zotero integration
  - Sharing Zotero files
  - Creating Zotero profiles
- Integrating Zotero into Obsidian (time permitting)

## Lecture Notes

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/02_1-slides")
slide_tabs(slide_details, "/slides/day2_1.html")
```

:::{.callout-tip}
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
Use right arrow <kbd>-></kbd> to scroll through slides
:::


## Videos

## In-class stuff

Here are all the materials we'll use in class:

Hands-on R materials:

- Reading files from the web

### 📦 Additional Resources

- [Click here](https://www.andrewheiss.com/blog/2021/12/01/multilevel-models-panel-data-guide/) for an example of how to do Bayesian multilevel models.
- Read Chapters 1 & 2 in [R for Data Science](https://github.com/bigdataforsocialjustice/PHHBHP7534/blob/main/reading/R_data_Science_partI.pdf)

### Competencies and Knowledge
- Foundational knowledge: 3, 6
- Competencies: 9
- MPH Specializations: 4
- PhD Specializations: 4

### Student Evaluation Activity for Assessment
- Exam 1
- Proposal 1

