---
title: "Associations and Confounding"
date: "2024-01-16"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sHjKJRcD-ebQUbcaFhc0cQH"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "S4OmWxc3h3Y", "Lecture Slides", "1", TRUE
 # "DGbLEgL4g6Y", "Drawing lines", "drawing-lines", TRUE,
#  "n-rV1TIoIgw", "Lines, Greek, and regression", "lines-greek-regression", FALSE,
 # "RYKsrTKWCR4", "Null worlds and statistical significance", "significance", FALSE
)
```

::: callout-note
Course objective(s) for this module: *Describe and execute the key steps
in designing, conducting, and evaluating research in health behavior and
health promotion.*
:::

## Readings

- {{< fa book >}} Howards, P. P. (2018a). An overview of confounding. Part 1: The concept and how to address it. Acta Obstetricia et Gynecologica Scandinavica, 97(4), 394–399. https://doi.org/10.1111/aogs.13295 [@howards_overview_2018]
- {{< fa book >}} Howards, P. P. (2018b). An overview of confounding. Part 2: How to identify it and special situations. Acta Obstetricia et Gynecologica Scandinavica, 97(4), 400–406. https://doi.org/10.1111/aogs.13293 [@howards_overview_2018-1]
- {{< fa book >}} Chapter 3 in [Impact Evaluation in Practice.](https://openknowledge.worldbank.org/server/api/core/bitstreams/4659ef23-61ff-5df7-9b4e-89fda12b074d/content) [@gertler_impact_nodate]
- {{< fa book >}} Chapter 7 *Telling the Story, Part II: Examining Factors that Measure the Reduction of Inequities and Disparities* in Cornelius, L. J., & Harrington, D. (2014). [A social justice approach to survey design and analysis.](https://books.google.com/books?hl=en&lr=&id=RFtuAwAAQBAJ&oi=fnd&pg=PP1&dq=social+justice+approach+to+survey+design&ots=DdB3oebcUd&sig=lTdyCQLtP0-pzWAaTk8R9H3BVGE) Pocket Guide to Social Work Research. 


### Your tasks to complete today

-   [ ] Select a research topic 

### Your tasks this week to keep you on track

- Conduct a literature search to identify articles for your literature review
- Create a new Zotero collection and begin downloading the articles into that collection

### In-class activities
- Conducting a literature review
- Creating bibliographies in Zotero


## Lecture Notes

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/day2")
slide_tabs(slide_details, "/slides/day2")
```

:::{.callout-tip}
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos

## In-class stuff

Here are all the materials we'll use in class:

Hands-on R materials:

- [R code](/example/Confounding.qmd) *Overview of Confounding, Part I*: Reproducing Howards' 
- [R code](/example/infant_mortality.qmd)  *Risk and Odds Ratios*: Measuring disparities and inequities
- [R code](/example/bias.qmd)  *Statistical Noise*: Random measurement error

### 📦 Additional Resources


### Competencies and Knowledge
- Foundational knowledge: 3, 6
- Competencies: 9
- MPH Specializations: 4, 7
- PhD Specializations: 4

### Student Evaluation Activity for Assessment
- Exam 1
- Proposal 1

