---
title: "Research Questions & Hypotheses"
date: "2024-01-23"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sHwqEeiG5xb3zAHcQCGWZUC"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "S4OmWxc3h3Y", "Lecture Slides", "1", TRUE
 # "DGbLEgL4g6Y", "Drawing lines", "drawing-lines", TRUE,
#  "n-rV1TIoIgw", "Lines, Greek, and regression", "lines-greek-regression", FALSE,
 # "RYKsrTKWCR4", "Null worlds and statistical significance", "significance", FALSE
)
```

::: callout-note
Course objective(s) for this module: 

- *Describe and execute the key steps
in designing, conducting, and evaluating research in health behavior and
health promotion.*
- Prepare a research product
:::


## Readings

- {{< fa podcast >}} Chapter 7 in Creswell, J. W., & Creswell, J. D. (2017). [Research design: Qualitative, quantitative, and mixed methods approaches.](https://github.com/bigdataforsocialjustice/PHHBHP7534/blob/main/ch7.pdf) Sage publications. 
- {{< fa book >}} Chapter 7 in [Statistics for People Who *Think* They Hate Statistics](https://github.com/bigdataforsocialjustice/PHHBHP7534/blob/main/reading/Salkind%20-%20Statistics%20for%20People%20Who%20(Think%20They)%20Hate%20Statis.pdf)

### Required Reading for PhD Students (recommended for others)


## Lecture Notes

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/day5")
slide_tabs(slide_details, "/slides/day5.html")
```

::: {.callout-note icon=false}

## Pay Attention

The [annotated bibliography](/resource/annotated.qmd) is due next class period

:::

### Competencies and Knowledge
- Foundational knowledge: 3, 6
- Competencies: 9
- MPH Specializations: 4, 7
- PhD Specializations: 4

### Student Evaluation Activity for Assessment
- Exam 1
- Proposal 1
