---
title: "Theory and Concepts"
date: "2024-01-25"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sFJtz7egfUASfhhz6zuj7GZ"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "S4OmWxc3h3Y", "Lecture Slides", "1", TRUE
 # "DGbLEgL4g6Y", "Drawing lines", "drawing-lines", TRUE,
#  "n-rV1TIoIgw", "Lines, Greek, and regression", "lines-greek-regression", FALSE,
 # "RYKsrTKWCR4", "Null worlds and statistical significance", "significance", FALSE
)
```


## Readings

- {{< fa book >}} Chapter 2 in Crosby, R. A., DiClemente, R. J., & Salazar, L. F. (Eds.). (2006). [Research methods in health promotion](https://github.com/bigdataforsocialjustice/PHHBHP7534/blob/main/ch2_philosophy_of_science.pdf) (pp. xx, 450). Jossey-Bass.
 [@crosby_research_2006]
- {{< fa book >}} Earp, J. A., & Ennett, S. T. (1991). [Conceptual models for health education research and practice.](https://academic.oup.com/her/article/6/2/163/2798165?login=true) Health Education Research, 6(2), 163–171. [@earp_conceptual_1991]
- {{< fa book >}} Chapters 1, 2 and 7 in Freudenberg, N., Klitzman, S., & Saegert, S. (n.d.). [Urban Health and Society.](https://github.com/bigdataforsocialjustice/PHHBHP7534/blob/main/Urban%20Health%20and%20Society_Interdisciplinary%20Approaches%20to%20Research%20and%20Practice.pdf) 
- {{< fa book >}} Chapter 5 [*Using Public Health to Reform The Legal and Justice Response to Domestic Violence*](https://github.com/bigdataforsocialjustice/PHHBHP7534/blob/main/Reconsidering_Law_and_Policy_Debates_A_Public_Health_Perspective-3.pdf) in Culhane, J. G. (Ed.). (2010). Reconsidering Law and Policy Debates: A Public Health Perspective. Cambridge University Press.


### Your tasks to complete today

-   [ ] Email me the Annotated Bibliography by midnight tonight
-   [ ] The subject of the email must be "*Annotated Bibliography for PHHBHP7534**"
-   [ ] Make sure to email me a Word document of the assignment as an attachment and that you followed the instructions outlined [here](resource/annotated.qmd) to complete the assignment
-   [ ] Complete Problem Set 3

### Your tasks this week to keep you on track

- Begin writing literature review
- Start studying for [Exam 1](resource/exam1.qmd)

::: {.callout-note icon=false}

:::

## Lecture Notes

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/06-slides")
slide_tabs(slide_details, "/slides/day6.html")
```

:::{.callout-tip}
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## Videos


## In-class stuff

Read my article on *Reconsidering Adverse Child Experiences*

### Competencies and Knowledge
- Foundational knowledge: 9
- Competencies: 9
- MPH Specializations: 3, 4
- PhD Specializations: 3, 4

### Student Evaluation Activity for Assessment
- Exam 1
- Proposal 1

