---
title: "Measures + Scale Development"
date: "2024-02-01"
---

```{r slides-videos, echo=FALSE, include=FALSE}
source(here::here("R", "slide-things.R"))
source(here::here("R", "youtube-playlist.R"))

playlist_id <- "PLS6tnpTr39sFVrHdTifzlyS9dsJjhEhsF"

slide_details <- tibble::tribble(
  ~youtube_id,   ~title, ~slide, ~active,
  "S4OmWxc3h3Y", "Lecture Slides", "1", TRUE
 # "DGbLEgL4g6Y", "Drawing lines", "drawing-lines", TRUE,
#  "n-rV1TIoIgw", "Lines, Greek, and regression", "lines-greek-regression", FALSE,
 # "RYKsrTKWCR4", "Null worlds and statistical significance", "significance", FALSE
)
```

## Readings

### Measurement

- {{< fa book >}} Chapter 3 *Developing a Foundation for the Study* in Cornelius, L. J., & Harrington, D. (2014). [A social justice approach to survey design and analysis.](https://books.google.com/books?hl=en&lr=&id=RFtuAwAAQBAJ&oi=fnd&pg=PP1&dq=social+justice+approach+to+survey+design&ots=DdB3oebcUd&sig=lTdyCQLtP0-pzWAaTk8R9H3BVGE) Pocket Guide to Social Work Research. 
- {{< fa book >}} Barboza, G. E. (2018). [Latent Classes and Cumulative Impacts of Adverse Childhood Experiences.](https://journals.sagepub.com/doi/full/10.1177/1077559517736628?casa_token=7cXsY6LHIlsAAAAA:TC5ZvFsskmwYjam-Rfk0w0sNe2AcqSuEZyvxzLy0FAh1MZM1hPjCxyuJcFE4bEwdLSNRwKmU1nJbygg) Child Maltreatment, 23(2), 111–125. https://doi.org/10.1177/1077559517736628

### Required Reading for PhD Students (and recommended for others)

::: {.callout-note icon=false}

## Pay Attention

Exam 1 will be administered next class.

:::

### Your tasks to complete today

-   [ ] Make sure that you review the study guide for Exam 1 and are prepared to take the Exam next class 

## Lecture Notes

The slides for today's lesson are available online as an HTML file. Use the buttons below to open the slides either as an interactive website or as a static PDF (for printing or storing for later). You can also click in the slides below and navigate through them with your left and right arrow keys.

```{r show-slide-tabs, echo=FALSE, results="asis"}
slide_buttons("/slides/04-slides")
slide_tabs(slide_details, "/slides/day8.html")
```

:::{.callout-tip}
**Fun fact**: If you type <kbd>?</kbd> (or <kbd>shift</kbd> + <kbd>/</kbd>) while going through the slides, you can see a list of special slide-specific commands.
:::


## In-class stuff

Here are all the materials we'll use in class:

### 📦 Additional Resources


### Competencies and Knowledge
- Foundational knowledge: 3, 6
- Competencies: 2, 9
- MPH Specializations: 4, 7
- PhD Specializations: 4

### Student Evaluation Activity for Assessment
- Exam 1
- Proposal 1


### Advanced Latent Variable Models

I have used LVM extensively in my own work. Here are some examples:

- Latent Class Analysis
  - Barboza, G. E. (2015). The association between school exclusion, delinquency and subtypes of cyber-and F2F-victimizations: Identifying and predicting risk profiles and subtypes using latent class analysis. Child Abuse & Neglect, 39, 109–122.
  - Barboza, G. E. (2018). Latent Classes and Cumulative Impacts of Adverse Childhood Experiences. Child Maltreatment, 23(2), 111–125. https://doi.org/10.1177/1077559517736628

- Growth Mixture Models
  - Barboza, G. E. (2020). Child Maltreatment, Delinquent Behavior, and School Factors as Predictors of Depressive Symptoms from Adolescence to Adulthood: A Growth Mixture Model. Youth & Society, 52(1), 27–54. https://doi.org/10.1177/0044118X17721803
  - Barboza, G. E., & Schiamberg, L. (2021). Dual trajectories of parenting self‐efficacy and depressive symptoms in new, postpartum mothers and socioemotional adjustment in early childhood: A growth mixture model. Infant Mental Health Journal, 42(5), 636–654. https://doi.org/10.1002/imhj.21937

- Parallel Process Models
  - Barboza, G. E., & Dominguez, S. (2017). Longitudinal growth of post-traumatic stress and depressive symptoms following a child maltreatment allegation: An examination of violence exposure, family risk and placement type. Children and Youth Services Review, 81, 368–378.
Barboza, G. E., Dominguez, S., & Pinder, J. (2017). Trajectories of post-traumatic stress and externalizing psychopathology among maltreated foster care youth: A parallel process latent growth curve model. Child Abuse & Neglect, 72, 370–382.


